Spatial Analysis

### 2021-09-28 Scripts/spatial_analisis.Rmd
Re-made figures 3 and 4 with new ports data

# Get final df with index
grid_eez_sw_df <- as.data.frame(grid_eez_sw_sf) %>%
  select(state,index) %>%
  left_join(ne_grid,
            by = "index")

@@

@@

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_sf_fp) %>%
  select(index,landing_port,species_name) %>% 
  left_join(ne_grid,
            by = "index")

@@

grids <- my_path("D", "Spatial/grid_eez_fp_sf", name = "grid_eez_fp_df.csv", read = T) %>%
  mutate(
    spatial = "fp"
  ) %>% 
  bind_rows(
    my_path("D", "Spatial/grid_eez_sw_sf", name = "grid_eez_sw_df.csv", read = T)
  ) %>% 
  mutate(
    spatial = ifelse(is.na(spatial),"sw",spatial)
  )# %>% 
  # spread(spatial,index)

@@

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  filter(!is.na(state)) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(lat = order,
    order = letters[1:11])

@@

ggsave("interpolation_fig.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 8,
       height = 6)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
  #           by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)) %>% 
  spread(spatial,mean_per) %>% 
  mutate(
    difference = (fp-sw)/((fp+sw)/2)*100,
    # difference = fp-sw,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,15,5),
                              limits=(c(0,15))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage difference", alpha = 0.8,option = "C",
                              breaks = seq(-20,70,10),
                              limits=(c(-20,70))
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg_spp_spatialN.jpg",
  p,
  base_height = 14,
  base_width = 14
)

@@

### Paper figure: Aggregated Proportional change

@@

  ggsave(filename = "area_plot_avg_spp_sptl.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 12,
         height = 10
  )

>>>>

# Get final df with index
grid_eez_sw_df <- as.data.frame(grid_eez_sw_sf) %>%
  select(state,abrev,index) %>%
  left_join(ne_grid,
            by = "index")

@@

state_names <- grid_eez_sw_df %>% 
  select(state,abrev) %>% 
  distinct(.keep_all = T)

@@

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_sf_fp) %>%
  select(index,landing_port,abrev=landing_state,spp=species_name) %>% 
  left_join(ne_grid,
            by = "index") %>% 
  left_join(state_names)

@@

grids <- my_path("D", "Spatial/grid_eez_fp_sf", name = "grid_eez_fp_df.csv", read = T) %>%
  mutate(
    spatial = "fp"
  ) %>% 
  bind_rows(
    my_path("D", "Spatial/grid_eez_sw_sf", name = "grid_eez_sw_df.csv", read = T)
  ) %>% 
  select(-spp) %>% 
  mutate(
    spatial = ifelse(is.na(spatial),"sw",spatial)
  ) %>% 
  distinct(.keep_all = T)

@@

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  filter(!is.na(state)) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(lat = order,
    # order = letters[1:11],
    order = c("b","a",letters[3:11]),)

@@

ggsave("interpolation_fig.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 8,
       height = 6)
```

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)) %>% 
  spread(spatial,mean_per) %>% 
  # Set NA's to 0 because no data found on FP
  mutate(fp = replace_na(fp,0)) %>% 
  # View()
  mutate(
    # difference = ifelse(fp == 0,100, (fp-sw)/((fp+sw)/2)*100),
    difference = abs(fp-sw),
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,35,5),
                              limits=(c(0,35))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Difference (%)", alpha = 0.8,option = "C",
                              breaks = seq(0,20,5),
                              limits=(c(0,22)),
                              # labels = c("<-10",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

save_plot(
  path = my_path("R","Figures"),
  "proportion_chg.jpg",
  p,
  base_height = 14,
  base_width = 14
)

@@

### Aggregated Proportional change

@@

  ggsave(filename = "area_plot_avg_spp_sptl.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 14,
         height = 8
  )
